---
title: "C240424_1"
output: pdf_document
date: "2024-12-05"
author: "YOUSEF IBRAHIM GOMAA MAHMOUD MABROUK"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- ## Problem 1-1 -->
## Problem 1-1:

<!-- Printing the PDF is the answer. -->

<!-- ## Problem 1-2 -->
## Problem 1-2:

*Yousef Ibrahim Gomaa Mahmoud Mabrouk*

**Yousef Ibrahim Gomaa Mahmoud Mabrouk**

<https://37743.github.io>

# Categorical Data Analysis

## RMarkdown

The sum of 2 and 3 is: `r 2+3`

* arbitrary list
  + bullet point 1
  + bullet point 2
  + bullet point 3
  
<!-- ## Problem 1-3 -->
## Problem 1-3:

```{r}
data("iris")
head(iris)
mean_sepal <- mean(iris$Sepal.Length)
mean_petal <- mean(iris$Petal.Length)
mean_sepal
mean_petal
var_sepal <- var(iris$Sepal.Length)
var_petal <- var(iris$Petal.Length)
var_sepal
var_petal
x <- iris$Sepal.Length
y <- iris$Petal.Length
cor(x, y)
plot(x,y)
```

According to the graph illustrated above, and the output value of correlation (`r cor(x, y)`), the Sepal Length (x) and Petal Length (y) features are positively correlated. (Directly Proportional)